---
title: "reesehs_OriginalHomeworkCode_01"
author: "Reese Hotten-Somers"
date: "2023-09-17"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Challenge 1

I attempt to extract every third element but I found some trouble along the way.  
However, Jimmy helped me at the begining of class(thanks Jimmy!) by using this code:  

library(stringr)  
jimmy_ulysses <- gsub("[[:punct:]]", replacement = "", ulysses)  
jimmy_ulysses <- unliststr_split(jimmy_ulysses, pattern = " ")  
unlist(jimmy_ulysses)[seq(from = 1, to = 22, by = 3)]  

Using Jimmy's code I was able to redo my code and figure out the challenge:  
```{r, reesecode}
library(stringr)
ulysses <- "Stately, plump Buck Mulligan came from the stairhead, bearing a bowl of lather on which a mirror and a razor lay crossed."
reese_ulysses<- gsub("[[:punct:]]", replacement = "", ulysses)
#this takes out all punctuation and replaces it with a space
reese_ulysses
reese_ulysses<-str_split(reese_ulysses, " ")
#this separates the string and puts an extra space between each word
reese_ulysses
reese_ulysses<-unlist(reese_ulysses)[seq(from=1, to= 22, by= 3)]
#makes string into vector again and generates sequence from vector using every 3 words
reese_ulysses
```


## Challenge 2
```{r, challenge 2}
m <- matrix(data = 1:80, nrow = 8, ncol = 10, byrow = FALSE)
m
#creates matrix
x<-m[,c(2,3,6)]
x
#shows columns 2, 3, and 6
x<-m[c(6,8),]
x
#shows rows 6 and 8
```



## Challenge 3
```{r, array challenge}
a<- array(data = 400:1, dim= c(5,5,4,4))
#creates array with numbers decending from 400-1 with they dimensions of 5,5,4,4
a 
a[1,1,1,2]
#returns data at that point in the array
a[2,3,2,]
#returns row
a[1:5,1:5,3,3]
#returns columns/rows 1-5 of the array 3,3
```


## Challenge 4:
```{r, primate list}
Platyrrhini<- c("Cebidae", "Atelidae", "Pitheciidae")
Catarrhini<- c("Cercopithecidae", "Hylobatidae", "Hominidae")
Tarsioidea <-c("Tarsiidae")
Anthropoidea <- list(Catarrhini, Platyrrhini)
names(Anthropoidea)<- list("Catarrhini", "Platyrrhini")
Haplorhini<- list(Anthropoidea, Tarsioidea)
names(Haplorhini)<- list("Anthropoidea", "Tarsioidea")
Lemuroidea <- c("Cheirogaleidae", "Lepilemuridae", "Indriidae", "Lemuridae",
    "Daubentoniidae")
Lorisoidea <- c("Lorisidae", "Galagidae")
Strepsirhini <- list(Lemuroidea, Lorisoidea)
names(Strepsirhini) <- list("Lemuroidea", "Lorisoidea")
Primates<- list(Haplorhini, Strepsirhini)
names(Primates)<- list("Haplorhini", "Strepsirhini")
Primates
#this creates a organized list of primate taxonomy with subsets 
```



## Challenge 5:
```{r, matrix manipulation }
d1<- c(3, 0, 1 ,23, 1, 2, 33, 1, 1, 42, 0, 1, 41, 0, 2) 
#data for matrix
m3<- matrix(d1, nrow=5, ncol=3, byrow=TRUE) #creates matrix with 5 rows/3 colums using d1 as data
m3
df<- data.frame(m3) #converts matrix into data frame do it can be manipulated differently
df
as.logical(df$X2) #makes column 2 logical valued
df
as.factor(df$X3) #makes colum 3 factor valued
str(df)
```
