---
title: "reesehs_FinalHomeworkCode_01"
author: "Reese Hotten-Somers"
date: "2023-09-19"
output:
  rmdformats::downcute:
    self_contained: true
    thumbnails: true
    lightbox: true
    gallery: false
    highlight: tango
# sofia and I spent a while finding the correct way to set up this theme, but it is so cute!
---

```{r setup, include=FALSE}
library(rmdformats)
```
## Reflections:

### Challenges Faced:

1. The main challenge that I faced was challenge 1. I was hard to understand how to use the split_str function, even when I looked up the help document. On my own I was able to understand how to split the string into individual letters or even words but it required the help of Jimmy to finally understand how to not get an out of bounds error. I then took Jimmy's code(which did not compile or run) and simplified it into a way that I understood. 

2. For challenge 2 I struggled with the last part, where you have to take the data from matrix point 2,2 to 6,9. For some reason all the commas and brackets and parentheses became a lot and I ended up having to trial and error my way through to the answer. In the end, it was a good excersise in troubleshooting and figuring out how to subset corrected. 

3. The final challenge I faced was trying to figure out how to select certain columns in a data frame in challenge 5. I tried to subset it, but it did ever quite work but then I was able to look into the as.factor function and in one of the examples it used the $ sign as a connector. I then looked at the structure of the data frame and I learned that the column names used X(instert number) notation so I was able to select the correct column. 

4. Side note: trying to find how to make the downcute theme(for the YAML) work was really hard because Sofia and I could not find the instructions!

### Nicole's Peery Commentary:

- The first thing I noticed is that in Challenge 2, I immediately noticed that  you put the rows and columns you were subsetting into a vector with c() and that made it so efficient.
- I would just add to challenge 2 the third part which is "Given the matrix, m, extract the elements from row 2, column 2 to row 6, column 9 and assign them to the variable x".
-In Challenge 5 your use of df$X2 and df$X3 to extract the 2nd and 3rd column to change their class was really great. I used [] notation to subset and it felt really clunky. If you want to save those changes to the data frame so the class of the column changed permanently, you can try df <- as.logical(df$X2).
-I also appreciate your use of hash comments in your code - it made it super easy for me to locate what you were doing and why.

### Peer Commentary, How it helped:

1. Peer commentary from Nicole first of all reminded me to finish challenge 2! I had inadvertently skipped the last part of the challenge and Nicole kindly reminded me to finish it. 

2. In one of Nicole's comments she suggested that I assign as.logical(df$X2) to a new data frame however, when I tried it messed my whole system up. So from this I learned that not all peer comments worked but it was a really interesting idea that I think could work in the future

3. Being able to peer comment on someone else's code really helped me as well. It was intriguing to see how my classmate worked through the problems in totally different ways that I did. it gave me some ideas for future ways of manipulating data!


## Challenge 1

I attempt to extract every third element but I found some trouble along the way.  
However, Jimmy helped me at the begining of class(thanks Jimmy!) by using this code:  

library(stringr)  
jimmy_ulysses <- gsub("[[:punct:]]", replacement = "", ulysses)  
jimmy_ulysses <- unliststr_split(jimmy_ulysses, pattern = " ")  
unlist(jimmy_ulysses)[seq(from = 1, to = 22, by = 3)]  

Using Jimmy's code(above) I was able to redo my code and figure out the challenge:  
```{r, reesecode}
library(stringr)
ulysses <- "Stately, plump Buck Mulligan came from the stairhead, bearing a bowl of lather on which a mirror and a razor lay crossed."
#creates string 
reese_ulysses<- gsub("[[:punct:]]", replacement = "", ulysses)
#this takes out all punctuation and replaces it with a space
reese_ulysses
#displays what reese_ulysses is so I can see how it had been changed
reese_ulysses<-str_split(reese_ulysses, " ")
#this separates the string and puts an extra space between each word
reese_ulysses
reese_ulysses<-unlist(reese_ulysses)[seq(from=1, to= 22, by= 3)]
#makes string into vector again and generates sequence from vector using every 3 words
reese_ulysses
```


## Challenge 2
```{r, challenge 2}
m <- matrix(data = 1:80, nrow = 8, ncol = 10, byrow = FALSE)
#creates matrix
m
x<-m[,c(2,3,6)]
#shows columns 2, 3, and 6
x
x<-m[c(6,8),]
#shows rows 6 and 8
x
x <- m[2:6,2:9]
#creates subset of matrix m extracting numbers from row 2, column 2 in the matrix to row 6, column 9.
#I initally forgot to include this in my Rmd file, but Nicole graciously reminded me to do it! Thank you :))
x
```



## Challenge 3
```{r, array challenge}
a<- array(data = 400:1, dim= c(5,5,4,4)) #creates array with numbers decending from 400-1 with they dimensions of 5,5,4,4
a 
a[1,1,1,2] #returns data at that point in the array
a
a[2,3,2,] #returns row
a
a[1:5,1:5,3,3] #returns columns/rows 1-5 of the array 3,3
a
```


## Challenge 4:
```{r, primate list}
Platyrrhini<- c("Cebidae", "Atelidae", "Pitheciidae")
#creates vector containing the list of strings and assigns it to the variable Platyrrhini(this is repeated many times for different taxonomies)
Catarrhini<- c("Cercopithecidae", "Hylobatidae", "Hominidae")
Tarsioidea <-c("Tarsiidae")
Anthropoidea <- list(Catarrhini, Platyrrhini)
#creates a list out of the previously created variables and assigns the list to another variable(Anthropoidea)
names(Anthropoidea)<- list("Catarrhini", "Platyrrhini")
#This then creates a list of strings and assigns Athropoidea as a name as the variable to the list of strings(process repeated more below) 
Haplorhini<- list(Anthropoidea, Tarsioidea)
names(Haplorhini)<- list("Anthropoidea", "Tarsioidea")
Lemuroidea <- c("Cheirogaleidae", "Lepilemuridae", "Indriidae", "Lemuridae",
    "Daubentoniidae")
Lorisoidea <- c("Lorisidae", "Galagidae")
Strepsirhini <- list(Lemuroidea, Lorisoidea)
names(Strepsirhini) <- list("Lemuroidea", "Lorisoidea")
Primates<- list(Haplorhini, Strepsirhini)
names(Primates)<- list("Haplorhini", "Strepsirhini")
Primates
#this creates a organized list of primate taxonomy with subsets 
```



## Challenge 5:
```{r, matrix manipulation }
d1<- c(3, 0, 1 ,23, 1, 2, 33, 1, 1, 42, 0, 1, 41, 0, 2) #data for matrix
m3<- matrix(d1, nrow=5, ncol=3, byrow=TRUE) #creates matrix with 5 rows/3 colums using d1 as data
m3
df<- data.frame(m3) #converts matrix into data frame do it can be manipulated differently
df
as.logical(df$X2) #makes column 2 logical valued and saves it to a new data frame
df
as.factor(df$X3) #makes column 3 factor valued
str(df) #reviews the structure of the data frame, I am able to see how the second column is not logical values using 0 and 1(for false and true of course) and column 3 is assigned levels(factor values)
```

